---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Longitudinal_Win_Odds

<!-- badges: start -->
<!-- badges: end -->

This repository contains R code for the simulation study and application example of the paper: *Longitudinal extension of the win odds for ordinal repeated measurements*

## The *lwo* package

We have created an R package [lwo](https://github.com/Yongxi-Long/lwo) for implementing the proposed method, it can be installed from GitHub via

```{r,eval=FALSE}
# install.packages("pak")
pak::pak("Yongxi-Long/lwo")
```


## Simulation study

-   The file lwo_simulation_study.qmd documents the setup of the simulation scenarios and the evaluation of the simulation results.
-   lwo_simulation_study.qmd calls the LWO_simulation.R to run simulations.

## Application to SID-GBS

-   The file lwo_analysis_example showcases the application of the proposed method to historical trial data, the SID-GBS trial (Walgaard, C., Second intravenous immunoglobulin dose in patients with Guillain-Barré syndrome with poor prognosis (SID-GBS): a double-blind, randomised, placebo-controlled trial. The Lancet Neurology, 20(4), 275-283.)
-   The trial data is not public-available, so the script also provides a way to simulate trial data.